Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-13 chi2012.tex
a bit of massage

In this section, we touch on motivating work for DataPalette from three fields: database integration, peronal information management, and end-user mashups and toolkits.

The primary motivation for this work stems from field studies in personal information management, where letting users effectively consolidating and work with data from multiple sources has remained a challenge for over a decade. The problem of data fragmentation (\cite{Jones05towardsa}, for example, derives from an inability to effectively consolidate data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}). In our own previously conducted field study, (reference anonymised) we observed several different variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) that people devised to merge in data from external sources -- be them online, offline, or other people.  Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
%% TODO MORE HERE

%% don't like the beginning of this > 
It became clear that effectively coping with data heterogeneity would be a central focus to letting people effectively draw upon multiple, arbitrary sources of data when consulting the theoretical data integration literature. Alon Halvey describes hetereogeneity as an inevitable by-product of the distributed processes that are used to author it:

\begin{quote} 
[The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs. Hence, even if they model overlapping domains, they will model them in different ways. Differing structures are a byproduct of human nature ?people think differently from one another even when faced with  the same modeling goal. \cite{alonhalevy}
\end{quote}

@@

A number of prototype PIM systems have dmonstrated data integration capacities. For example, SEMEX \cite{semex} was a PIM environment that provided users a uniform query interface over the contents of their personal information environments, across heterogeneous files, folders and repositories.  Similarly, Haystack was a personal information manager that allowed users to manipulate and managing their data in a uniform, consolidated manner \cite{haystack}. Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.

@@

The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instances’ schema’s structural and logical definitions \cite{castano2006matching}.  In our case, this is problematic because information from the web does not always have a schema.  In real-world scenarios, semi-automatic ontology matching approaches are favoured over automatic because reconciling the differences between ontologies that were designed for different purposes is not always accurate.  Semi-automatic approaches use varying levels of support to aid a user, from prompting users to manually check flagged alignments, to suggesting possible alignments.  Allowing users to compare instances across heterogeneous data is powerful because users can combine and evaluate multiple sources.  In principle, automatic instance matching approaches would allow users to seamlessly browse this data, however for now and the foreseeable future such approaches cannot make use of personal knowledge and do not guarantee their alignments.  Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.

\section{Pre-study 1 - Understanding Heterogeneity for Everyday Tasks}
\subsection{Method}
Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.  We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).  This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant’s experiences, while collecting the same general areas of information.  Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.

\subsection{Results and Discussion}
The participant demographic consisted of 8 people, 7 of which were male and 1 female.  They were split evenly into two age groups of 18-25 and 26-32.  All but one participant used social network sites very regularly, and they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site’s updates.  Those that used social networking websites primarily use Facebook.  All of the participants used Twitter to listen to broadcasted tweets, and 6 out of 8 people regularly tweeted.

@@

When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.  Also, their friend’s preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.  Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
\subsection{Prestudy 1 Summary of Findings}
People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.  They felt that organising an event should be a process that changes over time and did not need rushing.  Three of the participants said that they would like a recommendation system for places to eat and activities. Another strong requirement was that it had to be ubiquitous so that everyone could use it because they wanted a single place to communicate with their friends.  While Facebook was the most popular social networking site for organising events, half of the participants felt that it was not the best solution and would prefer a collaborative environment, which could be wiki based.

\section{Prestudy 2 : Technical Challenges of Integrating the Data}

\subsection{Method}

\subsection{Results}

\subsection{Analysis}

\subsection{Prestudy 2 Summary of Findings}

>>>>

In this section, we touch on motivating work for DataPalette from three fields: peronal information management, data integration, and end-user mashups and toolkits.

The primary motivation for this work stems from field studies in personal information management, where letting users effectively consolidate and work with data from multiple sources has remained a challenge for over a decade. The problem of data fragmentation (\cite{Jones05towardsa}, for example, derives from an inability to effectively consolidate data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}). In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage heterogeneous data from disparate sources -- be them online, offline, or other people.  Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
%% TODO MORE HERE

%% don't like the beginning of this > 
Enabling PIM tools to supporting ad-hoc data integration has remained
difficult because of the complex, theoretically difficult nature of data hetereogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes. As described by Alon Halvey:

\begin{quote} 
  [The problem stems from the fact that] we are trying to integrate   data systems that were developed for slightly (or vastly) different business needs. Hence, even if they model overlapping domains, they will model them in different ways. Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modeling goal. \cite{alonhalevy}
\end{quote}

Yet, research prototypes of PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.  For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.  Similarly, Haystack, a personal information manager, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}. Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.

@@

@@

The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.  In our case, this is problematic because information from the web does not always have a schema.  In real-world scenarios, semi-automatic ontology matching approaches are favoured over automatic because reconciling the differences between ontologies that were designed for different purposes is not always accurate.  Semi-automatic approaches use varying levels of support to aid a user, from prompting users to manually check flagged alignments, to suggesting possible alignments.  Allowing users to compare instances across heterogeneous data is powerful because users can combine and evaluate multiple sources.  In principle, automatic instance matching approaches would allow users to seamlessly browse this data, however for now and the foreseeable future such approaches cannot make use of personal knowledge and do not guarantee their alignments.  Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.

\section{Pre-studies: Understanding the use and shape of personal data}

% little introductory ditty about why we're doing pre-studies 
% actually we may wish to combine discussion of both (!)

\subsection{1 - Understanding Heterogeneity in Everyday Tasks}
\subsubsection{Method}
Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.  We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).  This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant’s experiences, while collecting the same general areas of information.  Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.

\subsubsection{Results}
The participant demographic consisted of 8 people, 7 of which were male and 1 female.  They were split evenly into two age groups of 18-25 and 26-32.  All but one participant used social network sites very regularly, and they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site’s updates.  Those that used social networking websites primarily use Facebook.  All of the participants used Twitter to listen to broadcasted tweets, and 6 out of 8 people regularly tweeted.

@@

When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.  Also, their friend’s preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.  Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.

\subsubsection{Summary of Findings}
People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.  They felt that organising an event should be a process that changes over time and did not need rushing.  Three of the participants said that they would like a recommendation system for places to eat and activities. Another strong requirement was that it had to be ubiquitous so that everyone could use it because they wanted a single place to communicate with their friends.  While Facebook was the most popular social networking site for organising events, half of the participants felt that it was not the best solution and would prefer a collaborative environment, which could be wiki based.

\subsection{2 - Technical Challenges of Integrating the Data}

The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.

\subsubsection{Method}
We sought out APIs and data feeds of popular Web services and sites that were aimed at developers targeting consumption of the data by end-users.  Based on statistics collected We first selected 4 different categories

\subsubsection{Results}

\subsubsection{Summary of Findings}
